### 2024-09-27 paper/paper.qmd
Second update with suggestions taken from peer review.

---
title: "An Investigation of Subway Delays in Toronto: Insights from 2023 TTC Data"
subtitle: "Diving deep into subway delays"
author: 
  - Charlie Zhang
thanks: "Code and data are available at: https://github.com/Zqyyk11/Investigation-on-Toronto-Subway-Delay."
date: "2024-09-23"
date-format: long
abstract: "This report presents an investigation into subway delays in Toronto using data from the Toronto Transit Commission (TTC) for the year 2023. The analysis examines patterns of delays across various subway stations, lines, and days of the week, with the goal of identifying critical factors that contribute to subway delays. Key findings show that Monday experienced the highest number of delays, with the Bloor-Danforth (BD) line being the most affected. These insights provide a data-driven foundation for improving operational efficiency and minimizing service interruptions in the future."
format: pdf
fontsize: 12pt
toc: true
number-sections: true
bibliography: references.bib
---
```{r}
#| include: false
#| warning: false
#| message: false
library(ggplot2)
library(here)
library(knitr)
library(opendatatoronto)
library(tidyverse)
library(dplyr)
library(lubridate)

cleaned_data = read_csv(here::here('data/analysis_data/cleaned_subway_delay_data.csv'))

```
# Introduction

Public transit plays a vital role in the daily lives of millions of people in urban centers around the world. In Toronto, the subway system, operated by the Toronto Transit Commission (TTC), serves as a backbone for many commuters. However, delays in subway services can cause significant inconvenience and disrupt the city’s flow of traffic. In 2023, Toronto’s subway system faced numerous delays, which are the focus of this report.

This analysis uses a comprehensive dataset of subway delays from the TTC to investigate the nature and extent of service disruptions. Specifically, the report aims to answer the following questions:

-   Which days of the week are most prone to delays?
-   What are the most affected subway lines and stations?
-   What times of day experience the most delays?

By answering these questions, the report seeks to provide actionable insights for improving the reliability of Toronto’s subway system, reducing the frequency and duration of delays, and enhancing the overall commuting experience.

The findings in this report are based on data-driven analyses and visualizations, which highlight delay trends throughout the year. This investigation serves as a foundation for further research into the causes of delays and offers guidance for policy makers and transit authorities in addressing key operational challenges.

# Data {#sec-data}

The data utilized in this report is sourced from Open Data Toronto and was imported using the opendatatoronto library [@gelfand2022]. The specific dataset used for analyzing TTC subway delays in Toronto is the "TTC Subway & SRT Train Service Delay Data" dataset [@toronto2024. All data analysis was conducted using R [@r2023], with the assistance of several key packages, including tidyverse [@wickham2019], dplyr [@wickham2023dplyr],here [@mueller2020] janitor [@firke2023], ggplot2 [@wickham2016], knitr [@knitr], and lubridate [@grolemund2011].

To transform the raw data into a clean, analysis-ready format, several preprocessing steps were carried out using R. First, unnecessary columns that did not contribute to the analysis, were removed to clean the dataset. Using the dplyr package, specific columns were selected based on relevance to subway delays, such as date, time, station, line, and delay duration. The data was further cleaned by handling any missing or inconsistent values with the janitor package to ensure data integrity. Any outliers or erroneous entries were also identified and either corrected or removed as appropriate. These steps allowed for the creation of a clean dataset, which provided a solid foundation for the analysis below. The entire process was conducted through efficient R code, leveraging the functionality of packages like dplyr for filtering and selection, and janitor for data cleaning.

## Overview of the data

```{r,echo=FALSE}
#| label: "Toronto Subway Delays 2023 Data"
#| fig-cap: "Sample of the Toronto subway's delay timer"
#| echo: false

# PREVIEW OF THE DATA
cleaned_data[, 1:7] %>% head() %>% knitr::kable()
```

## Subway delays by day of the week
```{r,echo=FALSE}
library(ggplot2)

# Assuming your cleaned data is in 'subway_data' and it has a 'Day' column
day_delay_summary <- aggregate(cleaned_data$`Min Delay`, by=list(cleaned_data$Day), FUN=sum)

# Plotting the total delays by day of the week
ggplot(day_delay_summary, aes(x=Group.1, y=x)) +
  geom_bar(stat="identity", fill="skyblue") +
  labs(title="Total Delays by Day of the Week (in minutes)",
       x="Day of the Week", y="Total Delay (Minutes)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

The study of subway delays on days of the week uncovers a noticeable trend linked to how weekdays and weekends unfold in terms of commuter patterns and travel demand shifts within the city's transit system. Observations show that Mondays encounter the most substantial delays and also other weekdays like Thursdays and Fridays due to heightened rider activity during the typical workweek when individuals travel to their workplaces. The uptick in passenger traffic during peak hours exerts pressure on subway operations by causing trains and busier platforms along, with prolonged boarding processes that collectively contribute to amplified service disruptions. During peak hours when there is demand for subway services issues like mechanical problems or signal interruptions tend to cause bigger disruptions compared to weekends. A reason weekends experience fewer delays possibly because fewer people use the subway due to less frequent service and reduced ridership. With a number of commuters relying on the subway during weekends the system runs more smoothly with less strain resulting in better train operations. Moreover weekend schedules often incorporate planned maintenance activities that have impact since there are fewer passengers traveling. The difference, between weekdays and weekends shows how commuter traffic affects the reliability of subway services. Weekdays experience frequent and severe delays during rush hour demands compared to weekends when operations are lighter and services run more smoothly.

## Subway delays by subway line
```{r,echo=FALSE}
# Aggregating the data by subway line
line_delay_summary <- aggregate(cleaned_data$`Min Delay`, by=list(cleaned_data$Line), FUN=sum)

# Plotting the total delays by subway line
ggplot(line_delay_summary, aes(x=Group.1, y=x)) +
  geom_bar(stat="identity", fill="salmon") +
  labs(title="Total Delays by Subway Line (in minutes)",
       x="Subway Line", y="Total Delay (Minutes)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

The Yonge University (Y U) and Bloor Danforth (B D) subway routes encounter delays mainly because they are the most heavily used and vital transportation links in Toronto city center and surrounding regions. These subway lines pass through populated neighborhoods and major transit junctions like downtown Toronto serving as crucial lifelines for the millions of people who rely on them daily for their commutes. Specifically the Y U Line provides access to locations such, as Union Station, financial centers and popular spots leading to congestion during busy periods. The BD Line runs through the city connecting neighborhoods from east to west. Is essential for many commuters getting to work or school and carrying out their daily routines efficiently. The large number of passengers using these lines presents difficulties like longer boarding times and crowded platforms along with frequent train delays. Furthermore the substantial ridership on these routes amplifies the effects of disruptions as any delays have a cascading effect throughout the system given the heavy dependence, on these crucial lines. On the hand the Sheppard line (SRT) has fewer delays compared to others mainly because it caters to smaller communities and operates with less frequent service thus easing the burden, on infrastructure and operations.

## Subway delays by top 10 stations
```{r,echo=FALSE}
# Aggregating and selecting the top 10 stations with the most delays
station_delay_summary <- aggregate(cleaned_data$`Min Delay`, by=list(cleaned_data$Station), FUN=sum)
top_10_stations <- station_delay_summary[order(-station_delay_summary$x),][1:10,]

# Plotting total delays by top 10 stations
ggplot(top_10_stations, aes(x=Group.1, y=x)) +
  geom_bar(stat="identity", fill="lightgreen") +
  labs(title="Total Delays by Top 10 Stations (in minutes)",
       x="Station", y="Total Delay (Minutes)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

The top ten stations experiencing the delays are mainly major terminals and transfer points in Torontos subway system – with Kennedy BD Station leading the list due to its operational intricacies and heavy passenger traffic flow through connections with other lines, like the Scarborough RT and the Bloor Danforth line. Currently at Eglinton Station construction is underway for the Eglinton Crosstown LRT project causing disruptions that affect passenger flow and train operations resulting in increased delays. Kipling and Bloor stations are also facing delays as they serve as transfer points for various subway and bus routes, with many commuters switching lines or modes of transportation. The clustering of travelers at these locations results in crowded platforms and increases the likelihood of delays since even small interruptions can lead to a domino effect that impacts the reliability of the service provided there.This scenario occurs mainly in stations serving as central transit points due to their passenger flow and intricate operations along, with occasional infrastructure issues that collectively add up to prolonged delay times.

## Subway delays Peak and off peak hours
```{r,echo=FALSE}
# Convert the 'Time' column from character to proper time format using lubridate's hms function
cleaned_data$Time <- hms::as_hms(cleaned_data$Time)

# Define peak hours (7:00 - 9:00 and 16:00 - 19:00) as well as off-peak hours
cleaned_data <- cleaned_data %>%
  mutate(
    period = case_when(
      (hour(Time) >= 7 & hour(Time) < 9) | (hour(Time) >= 16 & hour(Time) < 19) ~ "Peak",
      TRUE ~ "Off-Peak"
    )
  )

# Summarize the total delay minutes for Peak and Off-Peak
delay_summary <- cleaned_data %>%
  group_by(period) %>%
  summarise(total_delay_minutes = sum(`Min Delay`, na.rm = TRUE))

# Plotting the total delay for Peak vs. Off-Peak
ggplot(delay_summary, aes(x = period, y = total_delay_minutes, fill = period)) +
  geom_bar(stat = "identity") +
  labs(title = "Total Subway Delay Minutes: Peak vs Off-Peak Hours",
       x = "Time Period",
       y = "Total Delay Minutes") +
  theme_minimal()

```

The examination of subway disruptions in Toronto indicates a contrast between busy and quieter times of the day when it comes to delays. With less crowded periods encountering much longer total delay durations compared to peak hours despite the higher number of passengers and operational demands during those busier times when the transit system seems to focus on efficiency and manages to have fewer delays than, during the less busy times. Delays are more common during quieter hours due to frequent services and maintenance work being carried out then to avoid disruptions during prime commute times also fewer resources being available could lead to longer delays when problems occur. It seems that the system handles high demand situations effectively but could possibly do better at reducing interruptions during quieter times to maintain service all day long.

# Results

The analysis of Toronto subway delays was conducted across four key aspects: delays by day of the week, subway lines, stations, and peak vs. off-peak hours. Each of these aspects offers unique insights into the operational challenges faced by the Toronto Transit Commission (TTC) in managing subway services.

Delays by Day of the Week: The analysis revealed that Mondays experience the highest total delay minutes, likely due to the surge of passengers at the start of the workweek. Both Thursday and Friday also show elevated delays, potentially reflecting increased traffic as commuters prepare for the weekend. Weekends, on the other hand, see fewer delays, which can be attributed to lower ridership and reduced service frequency.

Delays by Subway Line:The Yonge-University (YU) and Bloor-Danforth (BD) lines experienced the most delays. As two of the busiest lines serving major commuter hubs in downtown Toronto, they face greater operational challenges and are more susceptible to disruptions due to overcrowding and high service frequency. In contrast, smaller lines, like the Sheppard and Scarborough RT, showed fewer delays, likely due to lower ridership and fewer trains.

Delays by Subway Station: Kennedy Station on the Bloor-Danforth (BD) line had the highest recorded delays, followed by Eglinton and Kipling stations. As key transfer and terminal points, these stations handle large volumes of passengers, which can cause bottlenecks, particularly during rush hours. Eglinton Station, which is undergoing heavy construction due to transit expansion projects, is another hotspot for delays.

Peak vs. Off-Peak Hours:The final analysis compared peak and off-peak hours, revealing that off-peak periods experienced significantly higher total delay minutes. This might be due to reduced service frequency during off-peak hours, which makes each delay more impactful. Scheduled maintenance, often performed during off-peak hours, and fewer operational resources could also contribute to the higher delay times.

The overall analysis gives an overview of subway disruptions in Toronto for the year 2023.The frequency of delays is greatly impacted by the day of the week – particularly Mondays and the latter part of the workweek witness the highest delays.The Yonge-University and Bloor-Danforth lines face delays owing to their popularity and frequent train services. Specific stations like Kennedy and Eglinton serve as congestion points. Accumulate more delay minutes due to their vital positions in the transit system. During times when there are passengers traveling such as off hours or low traffic periods of the day there tends to be more delays in terms of total minutes being experienced due, to possibly having fewer trains running and maintenance work taking place. The thorough analysis offers an overview of subway disruptions in Toronto for the year 2023 with delays being significantly impacted by various factors and influences within the system. The analysis points out areas for operational enhancements to concentrate on such as improving off hours service and handling delays at busy stations like Kennedy and Eglinton to enhance the overall efficiency of the system better. Likewise tackling the issues on subway lines like YU and BD could lead to fewer disruptions especially during peak hours, with heavy passenger traffic.

# Discussion

The analysis of subway delays in Toronto for 2023 reveals several important insights into the performance of the city's subway system. Delays are more pronounced during the workweek, particularly on Mondays and Thursdays, suggesting that the operational strain is highest at the start and end of the week, likely due to increased commuter traffic. The Yonge-University (YU) and Bloor-Danforth (BD) lines experience the most delays, reflecting their roles as major transit arteries with high passenger volumes and frequent service demands. Stations like Kennedy and Eglinton, which are key transfer points and hubs of construction activity, are more prone to delays, pointing to infrastructure and operational challenges that exacerbate service interruptions.

The removal of zero-delay entries further emphasizes that certain stations and lines consistently face challenges, while others see minimal delays. This highlights potential areas for targeted improvements, particularly on high-traffic lines and during peak commuting days. To reduce delays, strategic investments in infrastructure and more efficient scheduling could mitigate the operational pressures observed on busy lines and at critical stations. In addition, proactive measures, such as real-time monitoring and passenger load management, could further improve service reliability. Overall, the data points to a need for a more flexible and robust transit system capable of handling fluctuations in demand while minimizing disruptions.

\newpage

# References